---
title: "Assignment 5"
author: "Kyle Walker"
date: "9/30/2019"
output: html_document
---

### Assignment 5. Data Visualization with ggplot2
#### Submission Instruction. You will need to submit on Blackboard, in the Assignment section, a link to your assignment on your Github Webpage. Following this to create a github webpage and posted a link to your assignment. This assignment works with the c2015 dataset.

```{r}
library(tidyverse)
library(readxl)
library(stringr)
c2015 <- read_excel("C:/Users/student/Documents/Senior Year/MATH 421/Assignment 2/c2015.xlsx")
```

#### 1. Clean the data for easy graphing. Do the follows to clean and reduce the size of the data. 
##### Remove all observations that have a cell being either (1) NA, (2)‘Unknown’, (3)‘Not Rep’, or (4)‘Not Reported’
##### Remove all observations that have a cell containing either (1)‘Unknown’, (2)‘Not Rep’, or (3)‘Not Reported’. For instance, observations with DRINKING variable being Unknown (Police Reported) will be removed.
##### Fix TRAV_SP and AGE (following previous assignments) so that they are both numerics.
##### Filter so that there are only drivers in the data
```{r}
#Fixing Sex
c2015$SEX[is.na(c2015$SEX)]<- "Female"
#Fixing Age
c2015$AGE[c2015$AGE == 'Less than 1'] <- "0"
c2015$AGE <- as.numeric(c2015$AGE)
c2015$AGE[is.na(c2015$AGE)] <- mean(c2015$AGE)
#Fixing Trav_Sp
c2015$TRAV_SP <- str_replace(c2015$TRAV_SP, " MPH", "")
c2015$TRAV_SP <- str_replace(c2015$TRAV_SP, "Not Rep", "")
c2015$TRAV_SP <- str_replace(c2015$TRAV_SP, "Unknown", "")
c2015$TRAV_SP <- as.numeric(c2015$TRAV_SP)
c2015 = c2015[!(is.na(c2015$TRAV_SP)),]
#Getting rid of missing data
c2015 = c2015 %>%
  filter_all(~!is.na(.))
c2015 = c2015 %>%
  filter_all(~!(.=="Unknown"))
c2015 = c2015 %>%
  filter_all(~!(.=="Not Rep"))
c2015 = c2015 %>%
  filter_all(~!(.=="Not Reported"))
c2015 = c2015 %>%
  filter_all(~!(.==str_detect(.,"Not Rep")))
c2015 = c2015 %>%
  filter_all(~!(.==str_detect(.,"Unknown")))
c2015 = c2015 %>%
  filter(SEAT_POS == 'Front Seat, Left Side')
```

#### 2. Use geom_point to plot AGE and TRAV_SP coloring by SEX.
```{r}
ggplot(c2015, aes(x=AGE, y=TRAV_SP, col=SEX)) + geom_point()
```

#### 3. There is overplotting in 2. Overplotting is when many points are duplicated on the graph. Use geom_jitter instead of geom_point for 2. to avoid overplotting.
```{r}
ggplot(c2015, aes(x=AGE, y=TRAV_SP, col=SEX)) + geom_jitter()
```

#### 4. Plot histograms of AGE, TRAV_SP with bins = 50.
```{r}
ggplot(c2015, aes(x=AGE)) + geom_histogram(bins=50)
ggplot(c2015, aes(x=TRAV_SP)) + geom_histogram(bins=50)
```

#### 5. Plot a histogram of AGE coloring (fill) by SEX.
```{r}
ggplot(c2015, aes(x=AGE, fill=SEX)) + geom_histogram()
```

#### 6. Using geom_density to plot estimated densities of AGE colored by SEX.
```{r}
ggplot(c2015, aes(AGE, col=SEX)) + geom_density()
```

#### 7. Plot estimated densities of TRAV_SP colored by INJ_SEV.
```{r}
ggplot(c2015, aes(TRAV_SP, col = INJ_SEV)) + geom_density()
```

#### 8. Plot estimated densities of TRAV_SP seperated (colored) by weekdays and weekends.
```{r}
c2015 <- c2015 %>% 
  mutate(Weekend = case_when(
    DAY_WEEK %in% c("Saturday", "Sunday") ~ "Weekend",
    TRUE ~ "Weekday")
  )
ggplot(c2015, aes(TRAV_SP, col= Weekend)) + geom_density()
```

#### 9. Implement geom_bar on MONTH. Implement geom_bar on MONTH filled by SEX
```{r}
ggplot(c2015, aes(x=MONTH)) + geom_bar()
ggplot(c2015, aes(x=MONTH, fill=SEX)) + geom_bar()
```

#### 10. Implement geom_bar on MONTH and SEX with position='dodge'
```{r}
ggplot(c2015, aes(x=MONTH)) + geom_bar(position = "dodge")
ggplot(c2015, aes(x=SEX)) + geom_bar(position = "dodge")
```

#### 11. Plot a bar chart of average speeds in months using geom_col
```{r}
avgMSpd <- c2015 %>% 
  group_by(MONTH) %>% 
   summarize(AvgSpd = mean(TRAV_SP))
avgMSpd
ggplot(avgMSpd, aes(x=MONTH, y=AvgSpd)) + geom_col()
```

#### 12. Plot a bar chart of average speeds in months using geom_bar
```{r}
ggplot(avgMSpd, aes(x=MONTH, y=AvgSpd)) + geom_bar(stat = 'identity')


```

#### 13. Plot a bar chart of average speeds in months filled by SEX
```{r}
SpdSex <- c2015 %>% 
  group_by(MONTH, SEX) %>% 
    summarize(AvgSpd = mean(TRAV_SP))
SpdSex
ggplot(SpdSex, aes(x=MONTH, y=AvgSpd, fill=SEX)) + geom_bar(stat= "identity")
```

#### 14. Plot a bar chart of average speeds in months colored by INJ_SEV
```{r}
SpdSev <- c2015 %>% 
  group_by(MONTH, INJ_SEV) %>% 
    summarize(AvgSpd = mean(TRAV_SP))
SpdSev
ggplot(SpdSev, aes(x=MONTH, y=AvgSpd, fill=INJ_SEV)) + geom_bar(stat= "identity")
```


#### 15. Refer to this link to have a similar following plot:

##### Horizontal axis is for (monthly) average speed
##### The vertical axis is for months
##### Color by two colors: one for above overall average speed and the other for below the avarage speed
##### The speed on the horizontal axis is standardized
```{r}
cum_mean <- mean(c2015$TRAV_SP)
cum_sd <- sd(c2015$TRAV_SP)
gph15 <- c2015 %>%
  group_by(MONTH) %>% 
     summarize(spd_z = round(((mean(TRAV_SP,na.rm=T)- cum_mean) / cum_sd),2)) %>% 
       mutate(type = ifelse(spd_z > 0 , "above", "below")) %>% 
         arrange(spd_z)

ggplot(gph15, aes(x=reorder(MONTH,spd_z), y=spd_z, label = spd_z)) +
  geom_bar(stat='identity', aes(fill = type, width=0.5)) +
    scale_fill_manual(name = "Average Speed",
                      labels = c("Above Average" , "Below Average"),
                      values = c("above" = "#00ba38", "below" = "#f8766d")) +
  labs(subtitle = "Normalized Average Speed per Month",
       title = "Diverging Bars",
       x= "Month",
       y= "Normalized Speed") +
  coord_flip()

```


#### 16. Refer to this link to have a similar following plot:
##### Horizontal Axis is for mean speed
##### Vertical Axis is for INJ_SEV
##### Color by SEX
##### The numbers of speed are shown in points.
```{r}
gph16 <- c2015 %>% 
  group_by(INJ_SEV, SEX) %>% 
    summarize(spd= round(mean(TRAV_SP),2))

ggplot(gph16, aes(x=reorder(INJ_SEV,spd), y=spd, label = spd)) +
  geom_point(stat='identity', aes(col = SEX), size=9) +
    scale_color_manual(name = "Sex",
                      labels = c("Male" , "Female"),
                      values = c("Male" = "blue", "Female" = "red")) +
  geom_text(color="white", size=2) +
  labs(subtitle = "Average Speed per Month",
       title = "Diverging Dot Plot",
       x= "Injury Severity",
       y= "Mean Speed") +
  coord_flip()
```


#### 18. Refer to this link to have a similar following plot:
##### Horizontal Axis is for speed
##### Vertical Axis is for DAY
##### Color by SEX
##### The should be a invisible vertical line seperating the two sexes.
```{r}
brks <- seq(-60,60,10)
lbls <- paste0(as.character(c(seq(60,0,-10), seq(10,60,10))), "mph")


c2015 %>% 
  group_by(DAY, SEX) %>% 
    summarize(avg= round((mean(TRAV_SP, na.rm=T)),2)) %>%
      mutate(avg = ifelse(SEX == "Female",-avg, avg)) %>%
        ggplot(aes(x=reorder(DAY, avg), y= avg, fill= SEX)) +
        geom_bar(stat='identity', width=0.6) +
        scale_y_continuous(breaks = brks,
                           labels = lbls) +
        coord_flip() +
        labs(title= "Speed by Day and Sex",
             y = "Average Speed",
             x = "Day of the Month") +
        theme(plot.title = element_text(hjust= 0.5),
              axis.ticks = element_blank())
```


#### 18-20. Generate three other interesting graphs from the dataset.
```{r}
library(ggfortify)
theme_set(theme_classic())

ts <- c2015 %>% 
  group_by(DAY) %>% 
    summarize(avgSpd = mean(TRAV_SP))

ggplot(ts, aes(x=DAY)) +
  geom_line(aes(y=avgSpd)) +
  labs(title= "Travel Speed") +
  theme(plot.title = element_text(hjust= 0.5))

```

```{r}
ggplot(c2015, aes(SEX, AGE)) +
  geom_boxplot(varwidth = T, fill = "light blue") +
  labs(title = "Age by Sex",
       x = "Sex",
       y= "Age")
```

```{r}
library(ggcorrplot)

c2015_n <- c2015 %>% 
  select_if(is.numeric) %>% 
    select(-"YEAR")

corr <- round(cor(c2015_n),1)

ggcorrplot(corr, hc.order = T,
       type = "lower",
       lab = T,
       lab_size = 3,
       method = "circle",
       colors = c("tomato2", "white", "springgreen3"),
       title = "Correlogram of c2015",
       ggtheme = theme_bw)
```

